---
title: "Exploratory Data Analysis (EDA) for Four Year Graduation"
author: "Vanessa Gonzalez"
date: "`r format(Sys.Date())`"
output:
  pdf_document: default
  html_document:
    keep_md: yes
---

## Data Preparation and EDA
First File to Run
Open Libraries
```{r}
library("ggplot2", lib.loc="/Library/Frameworks/R.framework/Versions/3.4/Resources/library")
library("graphics", lib.loc="/Library/Frameworks/R.framework/Versions/3.4/Resources/library")
library("gplots", lib.loc="/Library/Frameworks/R.framework/Versions/3.4/Resources/library")
library("dplyr")
library("lattice", lib.loc="/Library/Frameworks/R.framework/Versions/3.4/Resources/library")
library("data.table", lib.loc="/Library/Frameworks/R.framework/Versions/3.4/Resources/library")
library("datasets")
library("readxl")
library("stats")
```
Import data set
```{r}
setwd("~/REGIS/Practicum I/RStudio")
dataSet <- read_excel("./dataSet.xlsx")
#View(dataSet)
```
Transform data set into data frame
```{r}
dfDataSet <- as.data.frame(dataSet)
head(dfDataSet)
```
```{r}
head(dfDataSet)
```

```{r}
summary(dfDataSet)
```
Transform Variable's type to the appropriate data type
```{r}
dfDataSet[1] <- lapply(dfDataSet[1], as.integer)
dfDataSet[3] <- lapply(dfDataSet[3], as.numeric)
dfDataSet[8:24] <- lapply(dfDataSet[8:24], as.numeric)
dfDataSet[2] <- lapply(dfDataSet[2], as.factor)
dfDataSet[4:7] <- lapply(dfDataSet[4:7], as.factor)
str(dfDataSet)
```
```{r}
summary(dfDataSet)
```
Transform data frame into a table
```{r}
tbDataSet <- data.table(dfDataSet)
tbDataSet

```
Create a table for variable "GraduationStatus"
```{r}
tb <- table(dfDataSet$GraduationStatus)
tb

```
Create table with percentages of total
```{r}
tb.prop <- dfDataSet$GraduationStatus %>%
   table() %>%
   prop.table() %>% {. * 100} %>% 
   round(2)

tb.prop
```
Create a Pie chart
```{r}
pie(tb, main = "CS Students Status that Started 2008-2014", col = c("Light Blue","Navy Blue","Dark Orange"))
```
Create a Bar chart
```{r}
barplot(tb, col = "navy blue", ylim=c(0,400),main = "Computer Science Students who Started the Program (2008-2014)")
```

Create a Bar plot where the "graduated"" students are divided into CS and Other Major students
```{r}
ggplot(dfDataSet) + 
  geom_bar(aes(x = factor(GraduationStatus), fill = CsGrad), position = 'dodge') +
  xlab('Graduation Status') + ylab('Frequency')
```
Create a plot to represent the CS students that have graduated in 4 years
```{r}
ggplot(dfDataSet) + 
  geom_bar(aes(x = factor(GraduationStatus), fill = dfDataSet$"4YG"), position = 'dodge') +
  xlab('Graduation Status') + ylab('Frequency')
```
Create a table of numbers and percetages of students that graduate in 4 years
```{r}
tb4YG <- table(dfDataSet$"4YG")
tb4YGpercent <- prop.table(tb4YG)
tb4YGpercent
cbind(tb4YG,prop.table(tb4YG))
```
Create a Pie Chart for Four-year graduation rate
```{r}
pie(tb4YG, main = "CS Students that Graduated in Four Years", col = c("Light Blue","Navy Blue","Dark Orange"))
```
Create a Bar chart for Four-year graduation rate
```{r}
barplot(tb4YGpercent, col = "navy blue", ylim=c(0,1),main = "Four Year Graduation Rate for CS Students who Started on 2008-2014")
```
Create a table of numbers and percetages of students that graduate in 5 years
```{r}
tb5YG <- table(dfDataSet$"5YG")
tb5YGpercent <- prop.table(tb5YG)
tb5YGpercent
cbind(tb5YG,prop.table(tb5YG))
```
Create a Pie Chart for Five-year graduation rate
```{r}
pie(tb5YG, main = "CS Students that Graduated in Five Years", col = c("Light Blue","Navy Blue","Dark Orange"))
```
Create a Bar Chart for Five-year graduation rate
```{r}
barplot(tb5YGpercent, col = "lightblue", ylim=c(0,1),main = "Five Year Graduation Rate for CS Students who Started on 2008-2014")
```
Create a table of numbers and percetages of students that graduate in 6 years
```{r}
tb6YG <- table(dfDataSet$"6YG")
tb6YGpercent <- prop.table(tb6YG)
tb6YGpercent
cbind(tb6YG,prop.table(tb6YG))
```
Create a Pie Chart for Five-year graduation rate
```{r}
pie(tb6YG, main = "CS Students that Graduated in Six Years", col = c("Light Blue","Navy Blue","Dark Orange"))
```
Create a Bar Chart for Six-year graduation rate
```{r}
barplot(tb6YGpercent, col = "navy blue", ylim=c(0,1), main = "Six Year Graduation Rate for CS Students who Started on 2008-2014")
```



