---
title: "CS Last Course Analysis"
author: "Vanessa Gonzalez"
date: "`r format(Sys.Date())`"
output: html_notebook
---
CS Grad: 
Yes - CS Students
OtherMajor - Left CS
NG - Not Graduated

Data from 2008-2018 and excludes current students and students that graduated from CS.

Import Data Set and open libraries.
```{r}
library(readxl)
CourseDateData <- read_excel("CourseDateData.xlsx")

library(readxl)
CourseDateDataNoCS <- read_excel("CourseDateDataNoCS.xlsx")
```
Take a look at the data.
```{r}
summary(CourseDateDataNoCS)
```

```{r}
str(CourseDateDataNoCS)
```
Remove unecesary columns
```{r}
DataNoCS <- CourseDateDataNoCS[-(1:6)]
str(DataNoCS)
```
Transform data type to the appropriate type.
```{r}
DataNoCS$GraduationStatus <- as.factor(DataNoCS$GraduationStatus)
DataNoCS$CsGrad <- as.factor(DataNoCS$CsGrad)
str(DataNoCS)
```
Remove Math Courses to understand better when students leave the CS program.
```{r}
CourseDates <- DataNoCS[-(1:2)]
CourseDates <- CourseDates[-(6)]
CourseDates <- CourseDates[-(9)]
CourseDates <- CourseDates[-(10)]
CourseDates <- CourseDates[-(11:13)]
str(CourseDates)
```

Transform data set into data frame.
```{r}
dfCourseDates <- as.data.frame(CourseDates)
```
Replace NA with 0
```{r}
dfCourseDates[is.na(dfCourseDates)] <- 0
head(dfCourseDates)
```
Find highest level last class each student took.
```{r}
LastClass <- colnames(dfCourseDates)[max.col(dfCourseDates,ties.method="first")]
dfCourseDates$LastClass <- LastClass
head(dfCourseDates)
```

```{r}
summary(dfCourseDates)
```
Open additional libraries
```{r}
library(caret)
library(lattice)
library(ggplot2)
```

Create table from data frame.
```{r}
tbLastClass <- table(dfCourseDates$LastClass)
tbLastClass
```
Create a data frame from table for graphing.
```{r}
dfLastClass <- as.data.frame(tbLastClass)
dfLastClass
```
Graph frequency of last CS class taken by students that left the program.
```{r}
p<-ggplot(data=dfLastClass, aes(x= reorder(Var1, Freq),  y=Freq)) +
  geom_bar(stat="identity", fill="steelblue") +
  labs(title ="Last Highest CS Course Taken by Not Grad. and Other Major Students", x = "Course", y = "Frequency") +
  ylim(0,60) +
  theme(axis.text.x=element_text(angle=90, hjust=1)) +
  geom_text(aes(label = Freq), hjust=-.5 , position = position_dodge(width = 1),inherit.aes = TRUE) +
  coord_flip()
p
```



