---
title: "Regression Classification Trees and Random Forest"
author: "Vanessa Gonzalez"
date: "`r format(Sys.Date())`"
output:
  pdf_document: default
  html_notebook: default
---

Open libraries
```{r}
library("mlbench")
library("dplyr")
library("caret")
library("randomForest")
library("lattice")
library("ggplot2")
library("rpart")
library("e1071")
library("caret", lib.loc="/Library/Frameworks/R.framework/Versions/3.4/Resources/library")
```


##Partitions Creation
```{r}
head(dfDataSet4YGImpute)
```
The RandomForest library does not accept numbers or "_" in columng titles so the column titles need to be modified.
```{r}
colnames(dfDataSet4YGImpute) <- c("FourYG", "One.CSCI101","One.MATH111","Two.CSCI261","Two.MATH112","Two.MATH201","Three.CSCI262","Three.MATH213","Four.CSCI341","Four.CSCI358","Four.MATH225","Five.CSCI306","Five.CSCI403","Five.MATH332","Six.CSCI406","Seven.CSCI370","Eight.CSCI400","Eight.CSCI442")
head(dfDataSet4YGImpute)
```
Parrtitions created with 80% of data for training and 20% of data for testing.
```{r}
inTraining <- createDataPartition(dfDataSet4YGImpute$FourYG, p = 0.80, list = FALSE)
training <- dfDataSet4YGImpute[inTraining, ]
testing <- dfDataSet4YGImpute[-inTraining, ]
```
```{r}
training
```

```{r}
testing
```

##Regresion Partition with method "class".
```{r}
FourYG.rp = rpart(FourYG ~ ., data=training, method = "class")
FourYG.rp
```
Display CP table for Fitted Rpart Object.
The main variables used in this classification tree were: CSCI206, CSCI403, CSCI341, and CSCI406.
```{r}
printcp(FourYG.rp)
```
Plot CP.
```{r}
plotcp(FourYG.rp)
```
To look at the importance of variables in the regression partition with method "class".
```{r}
summary(FourYG.rp)
```
Plot of classification Tree
```{r}
plot(FourYG.rp, uniform=TRUE, branch=.3, margin=0.2)
text(FourYG.rp, all=TRUE, use.n = TRUE)
```
Utilizing the regression classification tree model the results for the testing set are below.
```{r}
predictions = predict(FourYG.rp, testing, type="class")
table(testing$FourYG, predictions)

```

A confusion matrix is created to compare prediction results with testing results.
```{r}
library(caret)
confusionMatrix(table(predictions, testing$FourYG))
```

```{r}
min(FourYG.rp$cptable[,"xerror"])
```
```{r}
which.min(FourYG.rp$cptable[,"xerror"])
```
Prune the tree to increase accuracy.
Get the cost complecity parameter of the record
```{r}
FourYG.cp = FourYG.rp$cptable[3,"CP"]
FourYG.cp

```
```{r}
prune.tree = prune(FourYG.rp, cp= FourYG.cp)
```
```{r}
plot(prune.tree, margin= 0.1)
text(prune.tree, all=TRUE , use.n=TRUE)
```
Prune tree.
```{r}
prune.tree = prune(FourYG.rp, cp = FourYG.cp)
predictions.prune = predict(prune.tree, testing, type="class")
table(testing$FourYG, predictions.prune)

```
Confusion matrix for punned tree.
```{r}
confusionMatrix(table(predictions.prune, testing$FourYG))
```
## Random Forest Method

```{r}
FourYG.rf <- randomForest(FourYG ~One.CSCI101+One.MATH111+Two.CSCI261+Two.MATH112+Two.MATH201+Three.CSCI262+Three.MATH213+Four.CSCI341+Four.CSCI358+Four.MATH225+Five.CSCI306+Five.CSCI403+Five.MATH332+Six.CSCI406+Seven.CSCI370+Eight.CSCI400+Eight.CSCI442  , data = training)
FourYG.rf
```

```{r}
FourYG.rf.prediction <- predict(FourYG.rf, testing)
table(FourYG.rf.prediction, testing$FourYG)
```

To determine variable imortance.
```{r}
importance(FourYG.rf)
```
Plot for variable importance

```{r}
varImpPlot(FourYG.rf)
```


```{r}
confusionMatrix(table(FourYG.rf.prediction, testing$FourYG))
```
## Logistic Regression Method for Variable Importance
A different method was tried to confirm the variable importance obtained through Random Forest. The results were very similar.
```{r}
# Template code
# Build Logit Model on Training Dataset

FourYG.lr <- glm(FourYG ~One.CSCI101+One.MATH111+Two.CSCI261+Two.MATH112+Two.MATH201+Three.CSCI262+Three.MATH213+Four.CSCI341+Four.CSCI358+Four.MATH225+Five.CSCI306+Five.CSCI403+Five.MATH332+Six.CSCI406+Seven.CSCI370+Eight.CSCI400+Eight.CSCI442, family= "binomial", data = training)
FourYG.lr

# Predict Y on Test Dataset
predictedY <- predict(FourYG.lr, testing, type="response") 
```


Check prediction
```{r}
predictedY.rf <- predict(FourYG.rf, testing, type="response") 
predictedY.rf
```
Plot prediction
```{r}
plot(predictedY.rf, col = "navy blue")
```

For a list of importance of variables
```{r}
gbmImp <- varImp(FourYG.rf, scale = FALSE)
gbmImp
```


