---
title: "Model with Less Variables"
author: "Vanessa Gonzalez"
date: "`r format(Sys.Date())`"
output: html_notebook
---
## Random Forest Model using less variables

Open libraries.
```{r}
library("mlbench")
library("caret")
library("randomForest")
library("lattice")
library("ggplot2")
library("rpart")
library("e1071")
library("caret", lib.loc="/Library/Frameworks/R.framework/Versions/3.4/Resources/library")
```

```{r}
colnames(dfDataSet4YGImpute) <- c("FourYG", "One.CSCI101","One.MATH111","Two.CSCI261","Two.MATH112","Two.MATH201","Three.CSCI262","Three.MATH213","Four.CSCI341","Four.CSCI358","Four.MATH225","Five.CSCI306","Five.CSCI403","Five.MATH332","Six.CSCI406","Seven.CSCI370","Eight.CSCI400","Eight.CSCI442")
LessVariablesSet <- dfDataSet4YGImpute

# Creates Data Partitions and removes variables
inTrainingLess <- createDataPartition(LessVariablesSet$FourYG, p = 0.80, list = FALSE)
LessVariablesSet <- LessVariablesSet[-(2:3)]
LessVariablesSet <- LessVariablesSet[-(3)]
LessVariablesSet <- LessVariablesSet[-(5)]
LessVariablesSet <- LessVariablesSet[-(7)]
LessVariablesSet <- LessVariablesSet[-(11)]

# Creates Training data Set
trainingLess <- LessVariablesSet[inTrainingLess, ]
# Creates Testing data Set
testingLess <- LessVariablesSet[-inTrainingLess, ]
# Data Set with less varialbes
head(LessVariablesSet)
```
```{r}
trainingLess
```

```{r}
testingLess

```

##Regresion Partition with method "class" for set with less variables
```{r}
FourYG.rp.Less = rpart(FourYG ~ ., data=trainingLess, method = "class")
FourYG.rp.Less
```

```{r}

printcp(FourYG.rp.Less)
```
Create a summary for the data set with less variables
```{r}
summary(FourYG.rp.Less)
```
Prediction
```{r}
predictionsLess = predict(FourYG.rp.Less, testingLess, type="class")
table(testingLess$FourYG, predictionsLess)

```
Confusion Matrix
```{r}
library(caret)
confusionMatrix(table(predictionsLess, testingLess$FourYG))

```
```{r}
min(FourYG.rp.Less$cptable[,"xerror"])
```
```{r}
which.min(FourYG.rp.Less$cptable[,"xerror"])
```
Get the cost complecity parameter of the record
```{r}
FourYG.cp.Less = FourYG.rp$cptable[3,"CP"]
FourYG.cp.Less

```
```{r}
prune.tree.Less = prune(FourYG.rp.Less, cp= FourYG.cp.Less)
```


```{r}
prune.tree.Less = prune(FourYG.rp.Less, cp = FourYG.cp.Less)
predictionsLessPrune = predict(prune.tree.Less, testingLess, type="class")
table(testingLess$FourYG, predictionsLessPrune)

```

```{r}
confusionMatrix(table(predictionsLessPrune, testingLess$FourYG))
```

Top 10 variables

```{r}
str(LessVariablesSet)
```
## Random Forest model with les variables
```{r}
FourYG.rf.Less <- randomForest(FourYG ~Two.CSCI261+Two.MATH201+Four.CSCI341+Four.CSCI358+Five.CSCI306+Five.CSCI403+Five.MATH332+Six.CSCI406+Eight.CSCI400+Eight.CSCI442  , data = trainingLess)
FourYG.rf.Less
```

```{r}
FourYG.rf.prediction.Less <- predict(FourYG.rf.Less, testingLess)
table(FourYG.rf.prediction.Less, testingLess$FourYG)
```


```{r}
importance(FourYG.rf.Less)
```


```{r}
varImpPlot(FourYG.rf.Less)
```

```{r}
confusionMatrix(table(FourYG.rf.prediction.Less, testingLess$FourYG))
```
## Logistic Regresion with less variables

```{r}
# Template code
# Step 1: Build Logit Model on Training Dataset

FourYG.lr.Less <- glm(FourYG ~Two.CSCI261+Two.MATH201+Four.CSCI341+Four.CSCI358+Five.CSCI306+Five.CSCI403+Five.MATH332+Six.CSCI406+Eight.CSCI400+Eight.CSCI442, family= "binomial", data = trainingLess)
FourYG.lr.Less

# Step 2: Predict Y on Test Dataset
predicted.lr.Less <- predict(FourYG.lr.Less, testingLess, type="response") 
```
Variable Importance
```{r}
gbmImp.Less <- varImp(FourYG.rf.Less, scale = FALSE)
gbmImp.Less
```



