---
title: "Subsets of Data Set"
author: "Vanessa Gonzalez"
date: "`r format(Sys.Date())`"
output: html_notebook
---

## CS Graduated Students Data Set
```{r}
summary(dfDataSet)
```

Create a subset of data consisting of students with a "GraduationStatus" of "Graduated"
```{r}
GraduatedData<-subset(dfDataSet, GraduationStatus == 'Graduated')
head(GraduatedData)
```
Look at the data subset
```{r}
summary(GraduatedData)
```
```{r}
str(GraduatedData)
```


Remove not needed columns from data set and leave factor Four-year Graduation Factor
```{r}
DataSet4YG <- GraduatedData[(5:24)]
DataSet4YG <- DataSet4YG[-(2:3)]
head(DataSet4YG)
```

Transform data set into a data frame
```{r}
dfDataSet4YG <-as.data.frame(DataSet4YG)
```
Find coorelation between variables using "spearman" method
```{r}
res<- cor(dfDataSet4YG[-(1)], method = 'spearman', use = "complete.obs") 
round(res,2)
```

To substitute NA values with another value the KNN Imputation method is used
```{r}
library("DMwR")

DataSet4YGImpute <- knnImputation(DataSet4YG)
head(DataSet4YGImpute)
```
```{r}
str(DataSet4YGImpute)
```

```{r}
dfDataSet4YGImpute <-as.data.frame(DataSet4YGImpute)
```

Create a correlation plot between variables
```{r}
library(corrplot)
```
```{r}
corrplot(cor(dfDataSet4YGImpute[-(1)], method = 'spearman', use = "complete.obs"))
```
To determine variables with a correlation higher than 0.5
```{r}
highlyCorrelated <- findCorrelation(res, cutoff=0.5)
print(highlyCorrelated)
```

